### 2011-04-26 standing-patches-paper.tex
minor notes

Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
Therefore, if we define $\chi$ to be the unique positive solution to
\[
    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
\]
then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
before it is expected that one other successful mutation would have arisen in the area covered so far.
The explicit formula for $\chi$ is cumbersome but straightforward; here we omit the lengthy expression.
(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
This can be rewritten 
\begin{equation} \label{eqn:defines_chi}
    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^3 = (\pi \rho \mu)^{-1} .
\end{equation}
From this we see that $\chi$ increases linearly with $1/\sqrt{s_b}$,
and that the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
If $\sigma s_d \gg 1$, then standing variation is relatively unimportant, 
and this reduces to the characteristic length of \citet{ralphcoop2010}, $\chi_+ = (v/\lambda \pi)^{1/3}$.
On the other hand, if $\sigma s_d \ll 1$, then most adaptation come from standing variation, 
and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-2}$. 
% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.

@@

The quantity that is perhaps easiest to compute is the mean time until adaptation.
Fix some geographic location, and let $\tau\ge0$ be the time at which the point is reached by the advantageous mutations.
Then, as shown in Figure XX,
$\tau > t$ if and only if the cone with point at $(x,t)$ and slope $v$ extending back to $t=0$ is empty of successful mutations.
Since we assume these are a Poisson process, 
\[
    \P\{ \tau > t \} = \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 \right) ,
\]
and so we have
\begin{align}
    \E[\tau] % &= \int_0^\infty \P\{ \tau > t \} dt \\
        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 \right) dt.
\end{align}
For means of evaluting this integral, see Appendix \ref{apx:integrals}.

>>>>

Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
Therefore, if we define $\chi$ to be the unique positive solution to
\[
    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
\]
then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
before it is expected that one other successful mutation would have arisen in the area covered so far.
The explicit formula for $\chi$ is cumbersome but straightforward; here we omit the lengthy expression.
(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
This can be rewritten 
\begin{equation} \label{eqn:defines_chi}
    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^3 = (\pi \rho \mu)^{-1} .
\end{equation}
From this we see that $\chi$ increases linearly with $1/\sqrt{s_b}$,
and that the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
If $\sigma s_d \gg 1$, then standing variation is relatively unimportant, 
and this reduces to the characteristic length of \citet{ralphcoop2010}, $\chi_+ = (v/\lambda \pi)^{1/3}$.
On the other hand, if $\sigma s_d \ll 1$, then most adaptation come from standing variation, 
and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-1/2}$. 
\marginnote{XXX check this from below.}
% Note also substituting $z = \chi (\sqrt{2s_b} / (s_d \sigma)$ gets $z^2 + z^3 = (s_d^3 \sigma^2)/(\pi \rho \mu)$.
% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.

@@

The quantity that is perhaps easiest to compute is the mean time until adaptation.
Fix some geographic location, and let $\tau\ge0$ be the time at which the point is reached by the advantageous mutations.
Then, as shown in Figure XX,
$\tau > t$ if and only if the cone with point at $(x,t)$ and slope $v$ extending back to $t=0$ is empty of successful mutations.
Since we assume these are a Poisson process, 
\[
    \P\{ \tau > t \} = \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) ,
\]
and so we have
\begin{align}
    \E[\tau] % &= \int_0^\infty \P\{ \tau > t \} dt \\
        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) dt.
\end{align}
For means of evaluting this integral, see Appendix \ref{apx:integrals}.
